rgbif: a package for working with species occurrence data in R

### 2015-10-04 manuscript.Rmd
manuscript additions

Biodiversity loss is one of the greatest challenges of our
time [@Pimm_2014]. Some have called this the sixth great mass
extinction [@Ceballos_2015]. Given this challenge there is a
great need for, and much research is done on, biodiversity
science. The pool of questions about biodiversity we can answer
effectively becomes much larger with more and better data.

The Global Biodiversity Information Facility (hereafter, GBIF,
http://www.gbif.org/) is the largest collection of biodiversity
records globally, currently with 580 million records, 1.6 million
taxa, 15,000 datasets from 770 publishers. In addition, large
biodiversity warehouses such as iNaturalist, VertNet, and USGS's
BISON all feed into GBIF.

Herein, we describe a library (rgbif [@rgbif]) for working with
GBIF data in the R programming environment [@R]. R is an extremely
widely used language in academia, non-profits, and the private sector.
Importantly, R make it easy to do all of the steps of the
research process, including data management, data manipulation
and cleaning, statistics, and vizualization. An R client
for getting GBIF is a powerful tool for reproducible research.

# Install rgbif

@@

The `rgbif` package ...

@@

@@

>>>>

Perhaps the most fundamental element in many fields of ecology is
the inividual. How many individuals of each species in a given
location forms the basis for many sub-fields of ecology and evolution.
Some research questions necessitate collecting new data, while
others can easily take advantage of existing data. In fact, some
ecology fields are built largely on existing data, e.g., macro-ecology
[@Brown_1995; @Beck_2012].

Data on individuals, including which species, and where they're found,
can be used for a large number of research questions. In addition,
the pool of questions we can answer becomes much larger with more
and better data. In addition to wide utility, this data is important
for conservation. Biodiversity loss is one of the greatest challenges
of our time [@Pimm_2014]. Some have called this the sixth great mass
extinction [@Ceballos_2015]. Given this challenge there is a great
need for data on specimen records, whether collected from live sightings
in the field or specimens in museums.

There are many online services that collect and maintain specimen records.
However, Global Biodiversity Information Facility (hereafter, GBIF,
http://www.gbif.org/) is the largest collection of biodiversity
records globally, currently with 580 million records, 1.6 million
taxa, 15,000 datasets from 770 publishers (figures collected on 2015-10-04).
Many large biodiversity warehouses such as iNaturalist, VertNet, and USGS's
BISON all feed into GBIF.

Herein, we describe a library (rgbif [@rgbif]) for working with
GBIF data in the R programming environment [@R]. R is an extremely
widely used language in academia, and in non-profit and private sectors.
Importantly, R makes it easy to do all of the steps of the
research process, including data management, data manipulation
and cleaning, statistics, and vizualization. Thus, an R client
for getting GBIF data is a powerful tool for reproducible research.

@@

The `rgbif` package is completely written in R, uses an [MIT license][mit]
to maximize use everywhere. `rgbif` is developed publicly on GitHub
at [https://github.com/ropensci/rgbif][ghrgbif], where development versions
of the package can be installed, and bugs and feature requests reported.
Stable versions of `rgbif` can be installed from [CRAN][cranrgbif], the
distribution network for R packages. `rgbif` is part of the rOpenSci project,
a developer network making R software to facilitate reproducible research.

## Package interface

`rgbif` is designed following the [GBIF Application Programming Interface][gbifdocs],
or API. The GBIF API has four major components: registry, species names,
occurrence data, and maps. We ignore maps in `rgbif` as it is concerned with
generating maps for web applications. `rgbif` has a suite of functions dealing
with each of registry, species names, and occurrence data - we'll go through
each in turn describing design and example usage.

@@

The search API is very similar to the download API, but is meant for smaller
data acquisition jobs.

@@

[mit]: http://choosealicense.com/licenses/mit/
[ghrgbif]: https://github.com/ropensci/rgbif
[cranrgbif]: https://cran.rstudio.com/web/packages/rgbif/
[gbifdocs]: http://www.gbif.org/developer/summary
